---
title: "STAT530: HW2"
author: "Chunhua Ying"
date: "2020/2/9"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\textbf{Consider the data set website names} Hw1_2020.xls or .csv.  \textbf{Answer the questions based on this data set:}

\section{1}
\textbf{Plot the scatter plot and mention any possible violation of assumptions you see in this plot.}

From the plot: 1.the reationship is almost linear; 2.the variance is not increasing or decreasing obviously with respect to lead concentration; 3.No obvious outlier. Therefore no possible violation of assumption can be seen from this plot.
```{r}
#setwd("D:/Subject/WSU/STAT/Applied Linear Model/HW/HW2")
data2=read.csv("../HW_1_2020.csv")
plot(data2$leadconc,data2$mortality)
```





\section{2} 
\textbf{Plot the residual versus predicted plot and the normal probability plots.  What do you see?}

1. From residual versus predicted plot, the points are quite random, there is no obvious trend.

2. From the normal Q-Q plots, most of the residual data, especially in the middle, are along normal distribution. But the data at both left and right side would departure from Normality.

```{r}
data2=read.csv("../HW_1_2020.csv")
fit=lm(data2$mortality~data2$leadconc, data=data2)
summary(fit)
coefficients(fit)

#plot of residual vs fitted values
par(mfrow=c(1,2))
plot(data2$leadconc, fit$res)

# Normal probability plot of Residuals
qqnorm(fit$res)
qqline(fit$res)
hist(fit$res)
```





\section{3}
\textbf{Do the correlation test for normality.  For n>100, for alpha=.05, the critical value is: .987.}
Pvalue is smaller than 0.05, therefore reject H0(residuals are normally distributed)
```{r}
library(stats)
shapiro.test(fit$res)
```




\section{4}
\textbf{Do the Brown Forsyth Test or the Breusch Pagan test (whichever you think is appropriate).  Discuss results.}

Do the Breusch Pagan test below, the result Pvalue=0.8481, which is significantly larger than 0.05, so do not reject $H_0$(for log($\sigma_i$)=$\gamma_0+\gamma_1X_i$, $\gamma_1=0$), there is no sufficient evidence to conclude that variance linearly depend on X. This result support the assumption that the data set has equal variance.
```{r}
library(lmtest)
bptest(fit)
```


\section{5}
\textbf{Look at the Box-Cox transform to see what transformation if any is appropriate for the Y’s.}

From the plot below, The ML function for $\lambda$ would be maximum at a value close to -1. Therefore $\lambda=-1$ ($\frac{1}{\sqrt{Y}}$) would be appropriate transformation for Y's. 
```{r}
library(MASS)
boxcox(data2$mortality ~ data2$leadconc, data = data2,
       lambda = seq(-2, 2.0, length = 10))
```




\section{6}
\textbf{What is the value of the Durbin Watson statistic?}

DW=1.7537
```{r}
dwtest(fit)
```


\section{7}
\textbf{Give an example from your experience where you have faced multiple regression.}

The amount of money I would have to use each month versus electricity power used, Hours I stay outside, Hours that I spend on Amazon within the month. 



